Optimizing Abstract$^{\text 2}$ Machines

### 2012-11-11 paper/paper.tex
Toward more numbers.
More on store deltas.

The \emph{abstracting abstract machines}
approach~\cite{dvanhorn:VanHorn2011Abstracting,dvanhorn:VanHorn2012Systematic}
to deriving program analyses provides a straightforward way of
transforming a programming language semantics in the form of an
abstract machine, into a family of abstract interpreters parameterized
over policies for regulating the analytic precision of the control,
environment, store, and base value domains.  By taking a
machine-oriented view of computation, it becomes possible to design,
verify, and implement program analyzers for realistic language
features typically considered difficult to model.  The approach was
originally applied to features such as higher-order functions,
stack-inspection, exceptions, laziness, first-class continuations, and
garbage collection.  It has since been used to verify actor-
\cite{local:DOsualdo:12A} and
thread-based~\cite{dvanhorn:Might2011Family} parallelism and
behavioral contracts~\cite{dvanhorn:TobinHochstadt2012Higherorder}; it
has been used to model Coq~\cite{local:harvard},
Dalvik~\cite{local:dalvik}, Erlang~\cite{local:DOsualdo:12B},
JavaScript~\cite{local:DBLP:journals/corr/abs-1109-4467}, and
Racket~\cite{dvanhorn:TobinHochstadt2012Higherorder}.

The primary strength of the approach is that abstract interpreters can
be easily derived through a small number of steps from existing
machine models.  Since the relationships between abstract machines and
higher-level semantic models---such as definitional
interpreters~\cite{dvanhorn:reynolds-hosc98}, structured operational
semantics~\cite{dvanhorn:Plotkin1981Structural}, and reduction
semantics~\cite{dvanhorn:Felleisen2009Semantics}---are well
understood~\cite{dvanhorn:Danvy:DSc}, it is possible to navigate from
these high-level semantic models to sound program analyzers in a
systematic way.  Moreover, since these analyses so closely resemble a
language's interpreter (a) implementing an analysis requires little
more than implementing an interpreter, (b) a single implementation can
serve as both an interpreter and analyzer, and (c) verifying the
correctness of the implementation is straightforward.

However, there is a considerable weakness with the approach: an
analyzer designed and implemented by following the AAM recipe is
prohibitively inefficient without both further approximation and
further implementation effort.

In this article, we develop a systematic approach to deriving the
feasible implementation of an abstract machine based analyzer.

@@

\includegraphics[width=2.5in]{rel-time-church.ps}

@@

%% Notation: concrete examples of programs under analysis is given in
%% (monochrome) Scheme notation.  Code is given in (syntax colored)
%% Racket.

@@

\begin{figure}
\begin{alltt}
  ;; State \(\to\) \(\Set(\)State\()\)
  (define (step \(\mstate\))
    (match \(\mstate\)
      [(ev \(\mexp\) \(\rho\) \(\msto\) \(\mkont\))
       (match \(\mexp\)
         [(var\(\superscript\mlab\) \(\mvar\))
          (for/set ((\(\mval\) (lookup \(\rho\) \(\msto\) \(\mvar\))))
            (co \(\msto\) \(\mkont\) \(\mval\)))]
         [(lit\(\superscript\mlab\) \(\mlit\)) (set (co \(\msto\) \(\mkont\) \(\mlit\)))]
         [(lam\(\superscript\mlab\) \(\mvar\) \(\mexp\)) (set (co \(\msto\) \(\mkont\) (clos \(\mvar\) \(\mexp\) \(\rho\))))]
         [(app\(\superscript\mlab\) \(\mexp\subscript0\) \(\mexp\subscript1\))
          (define-values (\(\msto'\) \(\maddr\)) (push \(\mstate\)))
          (set (ev \(\mexp\subscript0\) \(\rho\) \(\msto'\) (ar \(\mexp\subscript1\) \(\rho\) \(\maddr\))))]
         [(ife\(\superscript\mlab\) \(\mexp\subscript0\) \(\mexp\subscript1\) \(\mexp\subscript2\))
          (define-values (\(\msto'\) \(\maddr\)) (push \(\mstate\)))
          (set (ev \(\mexp\subscript0\) \(\rho\) \(\msto'\) (ifk \(\mexp\subscript1\) \(\mexp\subscript2\) \(\rho\) \(\maddr\))))])]
      [(co \(\msto\) \(\mkont\) \(\mval\))
       (match \(\mkont\)
         ['mt (set (ans \(\msto\) \(\mval\)))]
         [(ar\(\superscript\mlab\) \(\mexp\) \(\rho\)) (set (ev \(\mexp\) \(\rho\) \(\msto\) (fn \(\mval\) l)))]
         [(fn\(\superscript\mlab\) \(\mval'\))
          (for/set ((\(\mkont\) (get-cont \(\msto\) l)))
            (ap \(\msto\) \(\mval'\) \(\mval\) \(\mkont\)))]
         [(fi\(\superscript\mlab\) c a \(\rho\))
          (for/set ((k (get-cont \(\msto\) l)))
            (ev (if v c \(\maddr\)) \(\rho\) \(\msto\) \(\mkont\)))])]
      [(ap \(\msto\) fun \(\maddr\) \(\mkont\))
       (match fun
         [(clos l \(\mvar\) \(\mexp\) \(\rho\))
          (define-values (\(\rho'\) \(\msto'\)) (bind \(\mstate\)))
          (set (ev \(\msto'\) \(\mexp\) \(\rho'\) \(\mkont\)))]
         [(op \(\mop\))
          (for*/set ((\(\mkont\) (get-cont \(\msto\) l))
                     (\(\mval\) (\(\interpdelta\) \(\mop\) \(\mval\))))
            (co \(\msto\) \(\mkont\) \(\mval\)))]
         [_ (set)]))]))
\end{alltt}
\caption{Implementation of machine transition relation.}
\end{figure}

@@

\begin{figure}
\begin{alltt}
(define (zero? n) ((n (\(\lambda\) (zx) #f)) #t))
(define (((mult m1) m2) mf) (m2 (m1 mf)))
(define ((((plus p1) p2) pf) x)
  ((p1 pf) ((p2 pf) x)))

@@

@@

;; multiplication distributes over addition
((church=? ((mult \church2) ((plus \church1) \church3)))
 ((plus ((mult \church2) \church1)) ((mult \church2) \church3)))
\end{alltt}
\caption{Church-numeral calculation}
\label{fig:church}
\end{figure}

Figure~\ref{fig:church} gives a small example of computing with Church
numerals.  The example, while small, is informative: (1) it can be
written in almost any modern programming language\footnote{It is
  written here using Scheme notation and makes use of the shorthands:
%
\begin{alltt}
   (define (((f x) y) z) ...) \(\equiv\)

   (define f (\(\lambda\) (x) (\(\lambda\) (y) (\(\lambda\) (z) ...))))
\end{alltt}

\begin{alltt}
   \church{n} \(\equiv\) (\(\lambda\) (s) (\(\lambda\) (z) (s\(\subscript1\) (s\(\subscript2\) ... (s\(\subscript{n}\) z)))))
\end{alltt}
},
%
and (2) it stresses an analyzer's ability to deal with complicated
environment and control structure arising from the use of higher-order
functions to model arithmetic.

Our baseline analyzer takes NN:NN minutes to analyze this program
(without store widening it times out after 30 minutes).  This is on
par with the performance of Sergey,~et al.'s analyzer, which takes
NN:NN minutes.

@@

The essence of the compilation effect can be seen by consider an example
such as
\[
\sapp{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2}}{\mexp_3}
\]
which makes the following transitions:
\begin{align}
& \ev{\sapp{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2}}{\mexp_3},\menv,\mkont,\msto_0}\\
\machstep\; &
\ev{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2},\menv,\kar{\mexp_3,\menv,\maddr_1},\msto_1}
\\
\machstep\; &
\ev{\sapp\mvar{\mexp_1},\menv,\kar{\mexp_2,\menv,\maddr_2},\msto_2}
\\
\machstep\; &
\ev{\mvar, \menv,\kar{\mexp_1,\menv,\maddr_3},\msto_3} % {\mexp_2}
\\
\machstep\; &
\co{\kar{\mexp_1,\menv},\mval,\msto_4} % {\mexp_1}{\mexp_2}
\mbox{ where } \mval \in \msto(\menv(\maddr))
\end{align}

@@

\begin{figure}
\begin{align*}
\mathit{eval}(\mexp) &= \{ \mstate\ |\ \compile{\mexp}(\epsilon,\varnothing,\varnothing,\kmt) \multimachstep \mstate \} \text{ where }
\\[2mm]
%% CONTINUE
\co{\kmt,\mval,\msto} &\machstep
\ans{\msto,\mval}
\\
\co{\kar[^\mcntr_\mlab]{\mcomp,\menv,\maddr},\mval,\msto} & \machstep
\mcomp^\mcntr(\menv,\msto,\kfn[^\mcntr_\mlab]{\mval,\maddr})
\\
\co{\kfn[^\mcntr_\mlab]{{\mvalx{u}},\maddr},\mval,\msto} & \machstep
\ap[^\mcntr_\mlab]{\mval,\mvalx{u},\mkont,\msto}
\text{ where }\mkont \in \msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\strue,\msto} & \machstep
\mcompi0^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\sfalse,\msto} & \machstep
\mcompi1^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\[2mm]
%% APPLY
\ap[^\mcntr_\mlab]{\clos{\mvar,\mcomp,\menv},\mval,\msto,\mkont} & \machstep
\mcomp^{\mcntr'}(\menv',\msto',\mkont) \\
\text{ where }\menv',\msto',\mcntr' &= \mathit{bind\,}^{ \mcntr}_\mlab(\msto,\mvar,\mval)
\\
\ap{\mop,\mval,\msto,\mkont} & \machstep
\co{\mkont,\mval',\msto} \\
\text{ where }\mkont &\in\msto(\maddr)
\text{ and } \mval'\in\interpdelta(\mop,\mval)
\end{align*}
\caption{Abstract$^2$ machine for compiled ISWIM}
\label{fig:caam}
\end{figure}

@@

Each step in the fixpointer for the above techniques requires joining
several (possibly large) stores together. Even with immutable hashes
that utilize sharing, this is memory-intensive. Worse, this
contributes an obscene fraction of the runtime.

\begin{alltt}
;; \(\widehat{State} \to \widehat{State}\)
(define (wide-step-specialized state)
  (match state
    [(cons \(\msto\) cs)
     (define-values (cs* \(\msdiff\))
       (for/fold ([cs* (set)] [\(\msdiff\)* '()])
         ([c cs])
         (match (step-compiled^ (cons \(\msto\) c))
           [(cons \(\msdiff\)** cs**)
            (values (\(\cup\) cs* cs**) (++ \(\msdiff\)** \(\msdiff\)*))])))
     (cons (update \(\msdiff\) \(\msto\)) (\(\cup\) cs cs*))]))
\end{alltt}

\[
\{ \widehat{v_3}, \widehat{v_5}, \widehat{v_6} \} = 0001011
\]

\[
\sigma = \mbox{\tt \#(0$_0$,\dots,0$_{|P|}$)}
\]

@@

The {\tt do} form has the special job of maintaining the current store
and accumulated states (if indeed those are being passed around). It
uses Racket's syntax parameters ~\cite{dvanhorn:eli/stxparam} to
track the most recent bindings of these accumulators so that {\tt
  yield-meaning} can refer to them (say {\tt $\sigma$p} and {\tt
  statep}). In the case of state passing and set-accumulated states, its
definition is simply

@@

\paragraph{Constraint-based program analysis for higher-order languages}

Constraint-based program analyses
(e.g.~\cite{dvanhorn:nielson-nielson-popl97,dvanhorn:wright-jagannathan-toplas98,dvanhorn:Meunier2006Modular})
typically compute sets of abstract values for each program point.
These values approximate the values arising at run-time for each
program point.  Value sets are computed as the least solution to a set
of constraints (either inclusion or equality constraints).  The
constraints must be designed and proved as a sound approximation of
the semantics.  Efficient implementations of these kinds of analyses
often take the form of worklist-based graph algorithms for contraint
solving, and are thus quite different from the interpreter
implementation.  The approach thus requires effort in constraint
system design and implementation, and the resulting system require
verification effort to prove the constraint system is sound and that
the implementation is correct.

This effort increases substantially as the comlexity of the analyzed
language increases.  Both the work of maintaining the concrete
semantics and constraint system (and the relations between them) must
be scaled simultaneously.  However, constraint systems, which have
been extensively studied in their own right, enjoy efficient
implementation techniques and can be expressed in declarative logic
languages that are heavily
optimized~\cite{dvanhorn:bravenboer-smaragdakis-oopsla09}.
Consequently, constraint-based analyses can be computed quickly.  For
example, Jagannathan and Wright's polymorphic splitting
implementation~\cite{dvanhorn:wright-jagannathan-toplas98} analyses
the church numeral example of figure~\ref{fig:church} in 7
milliseconds, about 25 times faster than the fastest implementation
considered here.  These analyses compute very different things (as
detailed in section~\ref{sec:accept}), so the performance comparison
is not apples-to-apples.  We view the primary contribution of this
work as a systematic path that eases the design, verification, and
implementation of analyses using the abstracting abstract machine
approach to within a factor of performant constraint-based analyses.

@@

%% \acks Sam Tobin-Hochstadt for encouragement and feedback -- he was
%% the first to prompt me to look into how make effective
%% implementations of the AAM approach.

%% NSF, DARPA

@@

\appendix
\section{Relation to Uniform \(k\)-CFA (A Case Against Acceptability)}
\label{sec:accept}

>>>>

The \emph{abstracting abstract machines}
approach~\cite{dvanhorn:VanHorn2011Abstracting,dvanhorn:VanHorn2012Systematic}
to deriving program analyses provides a straightforward way of
transforming a programming language semantics in the form of an
abstract machine, into a family of abstract interpreters parameterized
over policies for regulating the analytic precision of the control,
environment, store, and base value domains.  By taking a
machine-oriented view of computation, it becomes possible to design,
verify, and implement program analyzers for realistic language
features typically considered difficult to model.  The approach was
originally applied to features such as higher-order functions,
stack-inspection, exceptions, laziness, first-class continuations, and
garbage collection.  It has since been used to verify actor-
\cite{local:DOsualdo:12A} and
thread-based~\cite{dvanhorn:Might2011Family} parallelism and
behavioral contracts~\cite{dvanhorn:TobinHochstadt2012Higherorder}; it
has been used to model Dalvik, Erlang~\cite{local:DOsualdo:12B},
JavaScript~\cite{local:DBLP:journals/corr/abs-1109-4467},
Racket~\cite{dvanhorn:TobinHochstadt2012Higherorder}, and
Coq~\cite{local:harvard}.

The primary strength of the approach is that abstract interpreters can
be easily derived through a small number of steps from existing
machine models. % such as the CESK machine or the Lazy Krivine machine.
Moreover, since the relationships between abstract machines and
higher-level semantic models---such as definitional
interpreters~\cite{dvanhorn:reynolds-hosc98}, structured operational
semantics~\cite{dvanhorn:Plotkin1981Structural}, and reduction
semantics~\cite{dvanhorn:Felleisen2009Semantics}---are well
understood~\cite{dvanhorn:Danvy:DSc}, it is possible to navigate from
these semantic models to sound program analyzers.

@@

@@

Notation: concrete examples of programs under analysis is given in
(monochrome) Scheme notation.  Code is given in (syntax colored)
Racket.

@@

HERE: ρ

\begin{figure}
\begin{alltt}
  ;; State \(\to\) Set[State]
  (define (step \(\mstate\))
    (match \(\mstate\)
      [(ev \(\mexp\) \(\rho\) \(\msto\) \(\mkont\))
       (match \(\mexp\)
         [(var\(\superscript\mlab\) \(\mvar\))
          (for/set ((\(\mval\) (lookup \(\rho\) \(\msto\) \(\mvar\))))
            (co \(\msto\) \(\mkont\) \(\mval\)))]
         [(lit\(\superscript\mlab\) \(\mlit\)) (set (co \(\msto\) \(\mkont\) \(\mlit\)))]
         [(lam\(\superscript\mlab\) \(\mvar\) \(\mexp\)) (set (co \(\msto\) \(\mkont\) (clos \(\mvar\) \(\mexp\) \(\rho\))))]
         [(app\(\superscript\mlab\) \(\mexp\subscript0\) \(\mexp\subscript1\))
          (define-values (\(\msto'\) \(\maddr\)) (push \(\mstate\)))
          (set (ev \(\mexp\subscript0\) \(\rho\) \(\msto'\) (ar \(\mexp\subscript1\) \(\rho\) \(\maddr\))))]
         [(ife\(\superscript\mlab\) \(\mexp\subscript0\) \(\mexp\subscript1\) \(\mexp\subscript2\))
          (define-values (\(\msto'\) \(\maddr\)) (push \(\mstate\)))
          (set (ev \(\mexp\subscript0\) \(\rho\) \(\msto'\) (ifk \(\mexp\subscript1\) \(\mexp\subscript2\) \(\rho\) \(\maddr\))))])]
      [(co \(\msto\) \(\mkont\) \(\mval\))
       (match \(\mkont\)
         ['mt (set (ans \(\msto\) \(\mval\)))]
         [(ar\(\superscript\mlab\) \(\mexp\) \(\rho\)) (set (ev \(\mexp\) \(\rho\) \(\msto\) (fn \(\mval\) l)))]
         [(fn\(\superscript\mlab\) \(\mval'\))
          (for/set ((\(\mkont\) (get-cont \(\msto\) l)))
            (ap \(\msto\) \(\mval'\) \(\mval\) \(\mkont\)))]
         [(fi\(\superscript\mlab\) c a \(\rho\))
          (for/set ((k (get-cont \(\msto\) l)))
            (ev (if v c \(\maddr\)) \(\rho\) \(\msto\) \(\mkont\)))])]
      [(ap \(\msto\) fun \(\maddr\) \(\mkont\))
       (match fun
         [(clos l \(\mvar\) \(\mexp\) \(\rho\))
          (define-values (\(\rho'\) \(\msto'\)) (bind \(\mstate\)))
          (set (ev \(\msto'\) \(\mexp\) \(\rho'\) \(\mkont\)))]
         [(op \(\mop\))
          (for*/set ((\(\mkont\) (get-cont \(\msto\) l))
                     (\(\mval\) (\(\interpdelta\) \(\mop\) \(\mval\))))
            (co \(\msto\) \(\mkont\) \(\mval\)))]
         [_ (set)]))]))
\end{alltt}
\caption{Implementation of machine transition relation.}
\end{figure}

@@

%% \begin{alltt}
%% (define \church2 (\(\lambda\) (f) (\(\lambda\) (x) (f (f x)))))

%% (define pred
%%   (\(\lambda\) (n)
%%     (\(\lambda\) (rf)
%%       (\(\lambda\) (rx)
%%         (((n (\(\lambda\) (g) (\(\lambda\) (h) (h (g rf)))))
%%           (\(\lambda\) (i) rx))
%%          (\(\lambda\) (id) id))))))
%% \end{alltt}

%% (define \church0 (\(\lambda\) (f0) (\(\lambda\) (x0) x0)))
%% (define \church1 (\(\lambda\) (f1) (\(\lambda\) (x1) (f1 x1))))
%% (define \church2 (\(\lambda\) (f2) (\(\lambda\) (x2) (f2 (f2 x2)))))
%% (define \church3 (\(\lambda\) (f3) (\(\lambda\) (x3) (f3 (f3 (f3 x3))))))

\begin{figure}
\begin{alltt}
(define ((((plus p1) p2) pf) x)
  ((p1 pf) ((p2 pf) x)))

(define (((mult m1) m2) mf)
  (m2 (m1 mf)))

@@

(define (zero? n)
  ((n (\(\lambda\) (zx) #f)) #t))

@@

;; multiplication distributes over addition
((church=? ((mult \church2) ((plus \church1) \church3)))
 ((plus ((mult \church2) \church1)) ((mult \church2) \church3)))
\end{alltt}
\caption{Church-numeral calculation}
\end{figure}

Where \syntax{church=?} is an equality function for Church numerals
written in terms of recursion and an iterative \syntax{pred} function.
The \syntax{plus} and \syntax{mult} functions are written as usual.

This program reaches $N$ states under concrete interpretation.

@@

The essence of the compilation effect can be seen by considering an example
such as
\[
\sapp{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2}}{\mexp_3}
\]
which makes the following transitions:
\begin{align}
& \ev{\sapp{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2}}{\mexp_3},\menv,\mkont,\msto_0}\\
\machstep\; &
\ev{\sapp{\sapp\mvar{\mexp_1}}{\mexp_2},\menv,\kar{\mexp_3,\menv,\maddr_1},\msto_1}
\\
\machstep\; &
\ev{\sapp\mvar{\mexp_1},\menv,\kar{\mexp_2,\menv,\maddr_2},\msto_2}
\\
\machstep\; &
\ev{\mvar, \menv,\kar{\mexp_1,\menv,\maddr_3},\msto_3} % {\mexp_2}
\\
\machstep\; &
\co{\kar{\mexp_1,\menv},\mval,\msto_4} % {\mexp_1}{\mexp_2}
\mbox{ where } \mval \in \msto(\menv(\maddr))
\end{align}

@@

\begin{figure}
\begin{align*}
\mathit{eval}(\mexp) &= \{ \mstate\ |\ \compile{\mexp}(\epsilon,\varnothing,\varnothing,\kmt) \multimachstep \mstate \} \text{ where }
\\[2mm]
%% CONTINUE
\co{\kmt,\mval,\msto} &\machstep
\ans{\msto,\mval}
\\
\co{\kar[^\mcntr_\mlab]{\mcomp,\menv,\maddr},\mval,\msto} & \machstep
\mcomp^\mcntr(\menv,\msto,\kfn[^\mcntr_\mlab]{\mval,\maddr})
\\
\co{\kfn[^\mcntr_\mlab]{{\mvalx{u}},\maddr},\mval,\msto} & \machstep
\ap[^\mcntr_\mlab]{\mval,\mvalx{u},\mkont,\msto}
\text{ where }\mkont \in \msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\strue,\msto} & \machstep
\mcompi0^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\
\co{\kif[^\mcntr]{\mcompi0,\mcompi1,\menv,\maddr},\sfalse,\msto} & \machstep
\mcompi1^\mcntr(\menv,\msto,\mkont)
\text{ where }\mkont\in\msto(\maddr)
\\[2mm]
%% APPLY
\ap[^\mcntr_\mlab]{\clos{\mvar,\mcomp,\menv},\mval,\msto,\mkont} & \machstep
\mcomp^{\mcntr'}(\menv',\msto',\mkont) \\
\text{ where }\menv',\msto',\mcntr' &= \mathit{bind\,}^{ \mcntr}_\mlab(\menv,\msto,\mvar,\mval)
\\
\ap{\mop,\mval,\msto,\mkont} & \machstep
\co{\mkont,\mval',\msto} \\
\text{ where }\mkont &\in\msto(\maddr)
\text{ and } \mval'\in\interpdelta(\mop,\mval)
\end{align*}
\caption{Abstract$^2$ machine for compiled ISWIM}
\label{fig:caam}
\end{figure}

@@

Each step in the fixpointer for the above techniques requires joining
several, commonly large, stores together. Even with immutable hashes
that utilize sharing, this is memory-intensive. Worse, this
contributes an obscene fraction of the runtime. If we instead track
the changes to the store made over all steps and afterwards update the
store, we end up with much less allocation and computational overhead.

We represent changes as a list of pairs of address to set of storable
values. Each $\msto\sqcup[\maddr \mapsto \mval{s}]$ now becomes
$\cons{\ttuple{\maddr}{\mval{s}}}{\msdiff}$, where $\msdiff$ begins as empty for
each step. Applying the changes is straightforward:

\begin{align*}
\mtlst(\msto) &= \msto \\
\cons{\ttuple{\maddr}{\mval{s}}}{\msdiff}(\msto) &= \msdiff(\msto)\sqcup[\maddr \mapsto \mval{s}]
\end{align*}

The reduction relation does have to change to accomodate this
technique, but only in this regard. The whole store is still given so
that lookups are possible. Example rule:

\begin{align*}
bind^{\mcntr}_\mlab(\menv, \msto, \msdiff, \mvar, \mval) &= \menv[\mvar \mapsto \maddr], \cons{\ttuple{\maddr}{force(\msto,\mval)}}{\msdiff}, \mcntr' \\
\text{ where } \mcntr' &= \lfloor \mlab \mcntr\rfloor_k \\
               \maddr &= \mvar\mcntr' \\
(\ap[^\mcntr_\mlab]{\clos{\mvar,\mexp,\menv},\mval,\mkont},\msto,\msdiff) & \machstep
(\ev[^{\mcntr'}]{\mexp,\menv',\mkont},\msdiff') \\
\text{ where }\menv',\msdiff',\mcntr' &= \mathit{bind\,}^{ \mcntr}_\mlab(\menv,\msto,\msdiff,\mvar,\mval)
\end{align*}

We also lift $\machstep$ differently to accomodate for this asymmetry
in the input and output. For each state that is stepped, we feed the
output changes to the next so that all changes get accumulated.

\begin{align*}
(cs, \msto) &\widehat{\machstep} (cs\cup cs', \msdiff(\msto)) \\
\mbox{ where } (cs', \msdiff) &= step^*(\varnothing, cs, \mtlst) \\
step^*(S, \varnothing, \msdiff) &= (S, \msdiff) \\
step^*(S, \{c\}\cup cs, \msdiff &= step^*(S\cup cs^*, cs, \msdiff^*) \\
cs^* &= \{c' \mid (c,\msto,\msdiff) \machstep (c',\msdiff^c) \} \\
\msdiff^* &= concat\{\msdiff^c \mid (c, \msto,\msdiff) \machstep (c', \msdiff^c)\}
\end{align*}

@@

The {\tt do} form has the special job of maintaining the current store
and accumulated states (if indeed those are being passed around). It
uses Racket's syntax parameters ~\cite{ianjohnson:eli/stxparam} to
track the most recent bindings of these accumulators so that {\tt
  yield-meaning} can refer to them (say {\tt $\sigma$p} and {\tt
  statep}). In the case of state passing and set-accumulated states, its
definition is simply

@@

@@

@@

\appendix
\section{Relation to Uniform \(k\)-CFA (A Case Against Acceptability)}
